### 2016-03-22 talks/CausalBN/cbn2012.tex
x

\documentclass{beamer}
\usepackage{amsmath,amsfonts}	% use math symbols
\usepackage{graphicx} % insert images
\usepackage{url} % insert urls
\usepackage{textpos} % package for the positioning
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathtools}
\usetheme{Ilmenau} % change theme
%% NEW COMMANDS
\def\ci{\perp\!\!\!\perp}
\def\dep{\perp\!\!\!\perp\!\!\!\!\!\!\!/\,\,\,\,}
\begin{document}

%% MAKE TITLE
\title{Learning Causation from Data}
\subtitle{Fundamental of Casual Inference and its Applications}
\author{Huang Xiao} %\\ \small{xiaohu@in.tum.de} }%\thanks{Funded by ARAMiS project}}
%\and Author2 \inst{1} }
%\and Author3 \inst{2}}
\institute[Technische Universit\"at M\"unchen]
{ 
	%\inst{1} 
	Chair of IT Security (I20) \\Department of Informatics \\Technische Universit\"at M\"unchen 
	%\and
	%\inst{2} Cloud Service Lab\\Fraunhofer AISEC
}

\date{\today}
\maketitle

%% CHANGE THE DEFAULT TEMPLATE
% position the logo
\addtobeamertemplate{frametitle}{}{%
\begin{textblock*}{100mm}(0.98\textwidth,-0.7cm)
\includegraphics[height=0.53cm,width=0.9cm,keepaspectratio]{imgs/TUM}
\end{textblock*}}

\AtBeginSection[] {
	\frame<beamer>{ 
		\frametitle{Overview}   
		\tableofcontents[currentsection] 
 	}
 }
 % add frame number
%\setbeamertemplate{footline}{%
  %\raisebox{5pt}{\makebox[\paperwidth]{\hfill\makebox[10pt]{\scriptsize\insertframenumber}}}}

%% START PRESENTATION
\input{sections/fundamental}
\input{sections/causalBN}
\input{sections/copulaLearn}
\input{sections/experiments}

\section{Conclusion}
\begin{frame}{Wrap up}
\begin{itemize}
\item Causal Inference is a powerful tool for reasoning and predicting
\item Graphical model has a strong representation on causal conditions
\item Learning causation from data observation is totally feasible
\item Without additional information, only equivalence class can be achieved
\item Copula functions enable efficient parameter estimation
\item PICM based structure learning is also efficient
\item Promising experimental results show the interestingness of causal inference
\end{itemize}
\end{frame}

\section{References}
\begin{frame}{References}
\begin{thebibliography}{99}
\bibitem{Boas} J. Pearls,  Causality: Models, Reasoning, and Inference  \textit{Cambridge University Press}, \textbf{2000}.
\bibitem{Page} Joe Whittaker., Graphical Models in Applied Multivariate Statistics, \textit{John Wiley \& Sons}, Ltd, \textbf{2008}.
\bibitem{Page} Roger B. Nelsen, An Introduction to Copulas, \textit{Springer Science+Business Media}, Inc, \textbf{2006}.
\bibitem{Page} Jonas Peters, et al., Identifying Cause and Effect on Discrete Data using Additive Noise Models, \textit{JMLR}, Proceedings Track 9: 597-604, \textbf{2010}.
\end{thebibliography}
\end{frame}
\end{document}